---
title: "read_cdr_monthly"
authors: "Sam Kitrea"
date: "7/20/2021"
output: html_document
###July 20 2021
### SESYNC Sea Ice Stories
### Objective: Read in CDR Monthly data
### Output: Sea Ice Min Extent Time Series; Sea Ice Area Time Series

---
### Sea Ice Stories: SESYNC Workshop 2021
#### Identifying Sea Ice Anomalies
Samuel Aucoin, Kitrea Takata-Glushkoff


```{r setup, include=FALSE}
#Install Necessary Packages on First Run (comment out after first run)
#install.packages('ncdf4')

#Load Libraries
library(ncdf4) # package for netcdf manipulation
library(raster) # package for raster manipulation
library(rgdal) # package for geospatial analysis
library(ggplot2) # package for plotting
library(caTools)


#Settings
options(max.print=9999)
#Note - data file is in uploaded in this project.
#If you cannot access the data file within the project,
#set your filepath accordingly. 
getwd()

filepath <- '/Volumes/KTG_SSD/SESYNC/data/aggregate'

filename <- 'seaice_conc_monthly_nh_197811_202012_v04r00.nc'

#Load In Data
filename <- 'seaice_conc_monthly_nh_197811_202012_v04r00.nc'
nc_data <- nc_open(filename)

lon <- ncvar_get(nc_data, "longitude")
lat <- ncvar_get(nc_data, "latitude")
time <- ncvar_get(nc_data, "time")
sic <- ncvar_get(nc_data, 'cdr_seaice_conc_monthly')


sic[sic > 1] <- NaN                      #replacing flag values with NaNs

#Define Sea Ice Area
sia <- sic*(25^2)                        #sea ice area in each cell in km^2

#Define Sea Ice Extent
sie <- replace(sic, sic>=.15, 1)         #cells with sic >= 15 are considered ice-covered
sie <- replace(sie, sie<.15, 0)          #cells with sic < 15 considered ice-free
sie <- sie*(25^2)

#Aggregate cells across entire Arctic
icearea <- apply(sia, 3, sum, na.rm = TRUE)  #summing ice area across all pixels per month
total_sie <-  apply(sie, 3, sum, na.rm = TRUE) #not sure if this works yet

icearea[icearea == 0] <- NaN               #replacing months with missing data with NaNs
total_sie[total_sie == 0] <- NaN   


icearea <- icearea[c(-1, -2)]            #Deleting nov and dec of 1978
time <- time[c(-1, -2)]                  #Deleting nov and dec of 1978
total_sie <- total_sie[c(-1, -2)]        #Deleting nov and dec of 1978

yearly <- matrix(icearea, nrow = 42, byrow = TRUE) #each col is a month; rows are years

yearly_sie <- matrix(total_sie, nrow = 42, byrow = TRUE)


#Making the september sie anomaly

mean_sie <- mean(yearly_sie[3:32,9])     #Mean of September sie on the 1981-2010 period

anomaly_sie <- (yearly_sie[ , 9] - mean_sie)/1e+06

#or by percent
percent_anomaly <- (anomaly_sie/(mean_sie/1e+06))*100

t <- 1979:2020
sie_trend <- lm(formula = anomaly_sie ~ t)

#normalizing the data with the trendline
anomaly_norm <- anomaly_sie - sie_trend[["fitted.values"]]

#finding the standard deviation of the data
anomaly_sd <- sd(anomaly_norm)

```

### Figure 1.Sea Ice Area (September)
Region: Complete Arctic
Measure: Sea Ice Area in September of Each Year (from monthly average)


```{r, echo = FALSE }

### Plot September Total Sea Ice Area###
icearea_sep <- yearly[ , 9]              #taking september

icearea_sep <- icearea_sep/1e+06         #making units million km^2

plot(1979:2020, icearea_sep, type = 'b',
     main = 'September Arctic Sea Ice Area',
     xlab = 'Year',
     ylab = 'Ice Area (million km^2)',
     col = "darkblue",
     cex.lab = 1.2,
     cex.axis = 1, 
     font.lab = 1,
     cex = 0.7,
     pch = 16, 
     )

#abline(linregmod, col = "red", lty = 2) #add linear trendline
```


### Figure 2. September Sea Ice Extent
```{r, echo = FALSE }

### Plot ###
iceextent_sep <- yearly_sie[ , 9]         #taking september

iceextent_sep <- iceextent_sep/1e+06      #making units million km^2

plot(1979:2020, iceextent_sep, type = 'b',
     main = 'September Arctic Sea Ice Extent',
     xlab = 'Year',
     ylab = 'Ice Extent (million km^2)',
     col = "darkblue",
     cex.lab = 1.2,
     cex.axis = 1, 
     font.lab = 1,
     cex = 0.7,
     pch = 16,
     )

#abline(linregmod, col = "red", lty = 2)  #add linear trendline

```

### Figure 3. Arctic September Sea Ice Extent Anomaly (Area)
Region: Complete Arctic
Measure: September Sea Ice Extent Anomaly

```{r, echo=FALSE, fig.keep='all'}

### Plot September Total Sea Ice Extent Anomaly###

t <- 1979:2020
sie_trend <- lm(formula = anomaly_sie ~ t)

plot(1979:2020, anomaly_sie, type = 'b',
     main = 'September Arctic Sea Ice Extent Anomaly',
     xlab = 'Year',
     ylab = 'Ice Extent (million km^2)',
     col = "darkblue",
     cex.lab = 1.2,
     cex.axis = 1, 
     font.lab = 1,
     cex = 0.7,
     pch = 16,
     )+
  
  abline(sie_trend[["coefficients"]][["(Intercept)"]],sie_trend[["coefficients"]][["t"]], lty=6)+
  #adds 1 sigma lines
  abline(sie_trend[["coefficients"]][["(Intercept)"]]+anomaly_sd,sie_trend[["coefficients"]][["t"]], lty=6, col = 'darkgrey')+
  abline(sie_trend[["coefficients"]][["(Intercept)"]]-anomaly_sd,sie_trend[["coefficients"]][["t"]], lty=6,col = 'darkgrey')+

#adds 2 sigma lines
  abline(sie_trend[["coefficients"]][["(Intercept)"]]+2*anomaly_sd,sie_trend[["coefficients"]][["t"]], lty=3,col = 'darkgrey')+
  abline(sie_trend[["coefficients"]][["(Intercept)"]]-2*anomaly_sd,sie_trend[["coefficients"]][["t"]], lty=3,col = 'darkgrey')
  
legend('bottomleft', legend=c("trend", "1 stdev", "2 stdev"),col=c("black","darkgrey", "darkgrey"), lty=6:3, cex=0.8)

```

### Figure 4. Arctic September Sea Ice Extent Anomaly (%)

```{r, echo = FALSE }
percent_trend <- lm(formula = percent_anomaly ~ t)

plot(1979:2020, percent_anomaly, type = 'b',
     main = 'September Arctic Sea Ice Extent Anomaly',
     xlab = 'Year',
     ylab = 'Ice Extent Anomaly (%)',
     col = "darkblue",
     cex.lab = 1.2,
     cex.axis = 1, 
     font.lab = 1,
     cex = 0.7,
     pch = 16,
     )+
abline(percent_trend[["coefficients"]][["(Intercept)"]],percent_trend[["coefficients"]][["t"]])

#abline(linregmod, col = "red", lty = 2) #add linear trendline

```

### Figure 5. Monthly Sea Ice Extent with Moving Mean
Region: Complete Arctic
Measure: Monthly Sea Ice Extent with Moving Mean

```{r, echo = FALSE }
#Finding the moving mean

mov_sie <- runmean(total_sie,12)


plot(1:504, total_sie/1e+06, type = 'l',
     main = 'Arctic Sea Ice Extent',
     xlab = 'Year',
     ylab = 'Ice Area (million km^2)',
     col = "darkblue",
     cex.lab = 1.2,
     cex.axis = 1, 
     font.lab = 1,
     cex = 0.7,
     pch = 16,
     )+lines(1:504,mov_sie/1e+06,type = 'l')

```


